---
title: "Class 5: Data Visualization with ggplot"
author: "Yu-Chia Huang (PID: A59026739)"
format: pdf
---

## Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`.

Before I use any package functions I have to load them up with a `library()` call, like so:

```{r}
library(ggplot2)
#ggplot(cars)
```

```{r}
#cars
```

There is always the "base R" graphics system, i.e. `plot()`

```{r}
#plot(cars)
```

To use ggplot I need to spell out at least 3 things:

-   data (the stuff I want to plot as a data.frame)
-   aesthetics (aes() values - how the data map to the plot).
-   geoms (how I want things drawn)

```{r}
#ggplot(cars) +
  #aes(x=speed, y=dist) +
  #geom_point() +
  #geom_smooth(method="lm")
```

`geom_smooth()`using formula = 'y ~ x'


```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

Q. How many genes are there?

```{r}
nrow(genes)
```

There are `r nrow(genes)` in this dataset.

Q. What are the column names?
```{r}
colnames(genes)
```

The column names are `r colnames(genes)`.

Q. How many columns did you find?
```{r}
ncol(genes)
```

I found `r ncol(genes)` columns.

Q. How many are "up" regulated genes?

```{r}
sum(genes$State == "up")
```
There are `r sum(genes$State == "up")` up-regulated genes.

Q. What fraction of total genes is up-regulated in this dataset?
```{r}
round(table(genes$State)/nrow(genes) * 100, 2)
```

The fraction of up-regulated genes is `r round(sum(genes$State == "up")/nrow(genes) * 100, 2)`.

Q. Complete the code below to produce the following plot.

```{r}
p <- ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
p + scale_color_manual(values = c("cornflowerblue", "gray", "brown1")) +
  labs(title="Gene Expression Changes upon Drug Treatment") +
  xlab("Control (no drug)") +
  ylab("Drug Treatment")
```

