---
title: "Class 11 AlphaFold2 Analysis"
author: "Yu-Chia Huang (PID: A59026739)"
format: pdf
---

Here we post process and inspect our modeling results from AlphaFold2(AF).

My results from AF live in the folder/directory `hivprdimer_23119`

```{r}
results_dir <- "hivprdimer_23119/"

pdb_files <- list.files(results_dir, pattern = ".pdb", full.names = T)
basename(pdb_files)
```

We first need to align and supperpose these PDB models and we can use the `pdbaln()` function for this:

```{r}
library(bio3d)

# Read all data from Models 
#  and superpose/fit coords

pdbs <- pdbaln(pdb_files, fit=TRUE, exefile="msa")
```

```{r}
pdbs
```


## The RMSD matrix

A common measure of structural dis-similarity is called RMSD (root mean square distance).

```{r}
rd <- rmsd(pdbs, fit=T)
range(rd)
```

```{r}
#install.packages("pheatmap")
```


```{r}
library(pheatmap)

colnames(rd) <- paste0("m",1:5)
rownames(rd) <- paste0("m",1:5)
pheatmap(rd)
```

```{r}
# Read a reference PDB structure
pdb <- read.pdb("1hsg")
```


You could optionally obtain secondary structure from a call to `stride()` or `dssp()` on any of the model structures.

```{r}
plotb3(pdbs$b[1,], typ="l", lwd=2, sse=pdb)
points(pdbs$b[2,], typ="l", col="red")
points(pdbs$b[3,], typ="l", col="blue")
points(pdbs$b[4,], typ="l", col="darkgreen")
points(pdbs$b[5,], typ="l", col="orange")
abline(v=100, col="gray")
```


##Predicted Alignment Error for domains

Let's view these in Mol*. Here we want the fitted coords.

```{r}
xyz <- pdbfit(pdbs, core.inds, outpath="corefit_structures")
```

A full atom based fitting or supperposition did not work very well ecause we have multiple chains that are in different conformations.

I want to focus our supperposition on the most invarient part (the rigid "core" if you will).

```{r}
core <- core.find(pdbs)
```

```{r}
core.inds <- print(core, vol=0.5)
```

```{r}
xyz <- pdbfit(pdbs, core.inds, outpath="corefit_structures")
```

Now we can examine the RMSF between positions of the structure. RMSF is an often used measure of conformational variance along the structure:

```{r}
rf <- rmsf(xyz)

plotb3(rf, sse=pdb)
abline(v=100, col="gray", ylab="RMSF")
```

To evaluate how good multi-chain or multi-domain models are we need to look at the PAE scores (predicted aligned error).

There are output as JSON format files.

```{r}
#install.packages("jsonlite")
```


```{r}
library(jsonlite)

# Listing of all PAE JSON files
pae_files <- list.files(path=results_dir, pattern="0.json", full.names = TRUE)
pae_files
```

```{r}
pae1 <- read_json(pae_files[1], simplifyVector = TRUE)
pae5 <- read_json(pae_files[5], simplifyVector = TRUE)
```

```{r}
attributes(pae1)
```

```{r}
pae1$max_pae
pae5$max_pae
```

```{r}
plot.dmat(pae5$pae,
          xlab="Residue No.",
          ylab="Residue No.", zlim=c(0,30))
```

```{r}
plot.dmat(pae1$pae,
          xlab="Residue No.",
          ylab="Residue No.", zlim=c(0,16))
```

# Main points

We can run AlphaFold on google compute infrastructure :-)
We can read there results into R and process to help us make sense of these models and their PAE and pLDDT scores.

