---
title: "Lab 10 - Part 2"
author: "Yu-Chia Huang (PID: A59026739)"
format: html
---

Comparative Structure Analysis

Today we are going to finish off Lab 10 on learning about and analyzing protein structures.

##4. Comparative structure analysis of Adenylate Kinase

SEQ -> STRU -> FUNCTION

We will use the `bio3d` package for this analysis that starts with a single sequence.

We will also use the `msa` package from BioConductor.
First we need to install the `BiocManager`. We install it the same way as other packages with `install.packages()` compound.

We use `BiocManager::install()` to install any other BioConductor package we want - like `msa` in this case.


```{r}
library(bio3d)
```

```{r}
install.packages("BiocManager")
```

```{r}
library(BiocManager)
```

```{r}
BiocManager::install("msa")
```
```{r}
install.packages("devtools")
```

```{r}
devtools::install_bitbucket("Grantlab/bio3d-view")
```


First me will pick a sequence of interest - I will pick "1ake_A"

```{r}
aa <- get.seq("1ake_A")
aa
```

> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?

I want to search for all related structures in the PDB database.

```{r}
b <- blast.pdb(aa)
```

```{r}
#Plot a summary of search results
hits <- plot(b)
```


```{r}
head(b$hit.tbl)
```

```{r}
hits$pdb.id
```

> Side-note: Let;s save our results thaus far so we don't have to run blast again.

```{r}
save(hits, b, file="blast_results.Rds")
```

```{r}
load("blast_results.Rds")
hits
```

Now we will download all these related structures from the daatabase with `get.pdb()`.

```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
# Align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

##Principal Component Analysis

We can do a PCA on the coordinate data (ie.e the `$xyz` values) of all these structures.

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
dim(pdbs$xyz)
```

```{r}
plot(pc.xray, 1:2)
```

```{r}
mktrj(pc.xray, file="pca_results.pdb")
```





