---
title: "Class 10 Halloween"
author: "Yu-Chia Huang (PID: A59026739)"
format: pdf
---

##Importing candy data
```{r}
candy <- read.csv("candy-data.csv", row.names = 1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

There are `r nrow(candy)`.

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

```{r}
candy[as.logical(candy$fruity),]
```

##2. What is your favorate candy?

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Hershey's Kisses",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

##Side-note: the skimr::skim() function

#Use :: then we can do the same thing (install the package and call it).
#install.packages("skimr")
#library("skimr")

```{r}
skimr::skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Sure is, Winpercent.

>Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}
candy$chocolate
```
It represents whether each candy is chocolate (1) or not chocolate(0).

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

>Q9. Is the distribution of winpercent values symmetrical?

No. It's not symmetrical.

>Q10. Is the center of the distribution above or below 50%?

```{r}
choc.ind <- as.logical(candy$chocolate)
fruit.ind <- as.logical(candy$fruity)

choc.win <- candy[choc.ind,]$winpercent
fruit.win  <- candy[fruit.ind,]$winpercent

mean(candy$winpercent)
```
It's above 50%.

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(choc.win)
mean(fruit.win)
```
Chocolate is higher ranked than fruit candy.

>Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruit.win)
```

Yes. The p-value is 2.871e-08. We can reject the null hypothesis.

##3. Overall Candy Rankings
>Q13. What are the five least liked candy types in this set?

```{r}
x <- c(5,2,3,6)
sort(x)
sort(x, decreasing = T)
```

```{r}
x
order(x)
```

```{r}
x[order(x)]
```

```{r}
y <- c("D", "A", "E")
order(y)
```

```{r}
inds <- order(candy$winpercent)
head(candy[inds,], 5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
inds <- order(candy$winpercent, decreasing = T)
head(candy[inds,], 5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library("ggplot2")
ggplot(candy) +
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

##Time to add some useful color
Let's color up these candy up by some scheme.

```{r}
mycols <- rep("black", nrow(candy))
mycols[as.logical(candy$chocolate)] <- "chocolate"
mycols[as.logical(candy$bar)] <- "brown"
mycols[as.logical(candy$fruity)] <- "pink"

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col(bg=mycols)
```

Now, for the first time, using this plot we can answer questions like:

>Q17. What is the worst ranked chocolate candy?

Sixlets is the worst ranked chocolate candy.

>Q18. What is the best ranked fruity candy?

Starburst is the best ranked fruity candy.

##4. Taking a look at pricepercent

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?


```{r}
#install.packages("ggrepel")
```

```{r}
library("ggrepel")
library("ggplot2")
# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=mycols) + 
  geom_text_repel(col=mycols, size=3.3, max.overlaps = 5)
```

Reese's Miniatures.

```{r}
ord <- order(candy$pricepercent, decreasing = FALSE)
head( candy[ord,c(11,12)], n=5 )
```

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

Nik L Ni, Ring pop, Nestle Smarties, Pop Rocks, Sugar Babies.

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


Optional
>Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().



##5 Exploring the correlation structure
# Correlation structure

```{r}
library("corrplot")
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

>Q23. Similarly, what two variables are most positively correlated?

##6. Principal Component Analysis

```{r}
pca <- prcomp(candy,scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```

```{r}
plot(pca$x[,1:2], col=mycols, pch=16)
```
